---
title: "Bicycle Analysis"
author: "Todd Idol"
date: "10/10/2020"
output: 
   github_document:
    toc: true
    toc_depth: 2
params:
  day: 1
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
rmarkdown::find_pandoc(version = '2.10.1')
options(knitr.duplicate.label = "allow")
library(tidyverse)
library(jsonlite)
library(httr)
library(rmarkdown)
library(knitr)
library(caret)
```

# Project Repo

Find the project repo [here](https://github.com/tkidol/ST558-Project-2).

# Weekly Analysis

The analysis for [Sunday is available here](SundayAnalysis.md).
The analysis for [Monday is available here](MondayAnalysis.md).
The analysis for [Tuesday is available here](TuesdayAnalysis.md).
The analysis for [Wednesday is available here](Wedesdaynalysis.md).
The analysis for [Thursday is available here](ThursdayAnalysis.md).
The analysis for [Friday is available here](FridayAnalysis.md).
The analysis for [Saturday is available here](SaturdayAnalysis.md).

# Packages

## Tidyverse

The workhorse of this project for 3 main core packages: DPLYR - manipulated all of my data (selects, joins, filter, new variables... ), Tibble for rendering Data Frames more effectively, & () ggplot2 for discrete & continuous data plots

## Rmarkdown 

Key to the project for the Rmd file itself including the knitr:: functions for consumable object output (kable) & README.md github doc output through render()

## Caret

Used to create standardized train/test data & create controls and tuning for LM & GLM analysis

# Set Parameters & Knit

# Data

```{r, data}

bikeData <- read.csv("hour.csv")

# removing casual & registered vars since they will not be used for modling
bikeData <- bikeData  %>% select(everything(), -c("casual", "registered")) %>% mutate(season = as.factor(season), mnth= as.factor(mnth), hr = as.factor(hr), weekday = as.factor(weekday), workingday = as.factor(workingday), weathersit = as.factor(weathersit))

bikeData <- bikeData  %>% filter(weekday == params$day)

# Changing weekday variable from numeric to character

knitr::kable(head(bikeData))
```
# Create Parameters & Render

#{r, parameters}




